# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r echo = FALSE, message = FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data", "burden-of-disease-cmnn.csv"))

```

## Table of Estimates for CMNN Burden Over Time

```{r echo = FALSE, message = FALSE, show_col_types = FALSE}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

ben_mli_bfa_cmnn <- data_cmnn %>% 
  filter(Entity %in% c("Benin","Burkina Faso", "Mali")) %>% 
  select(Entity,Year,`DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  rename(DALYs = "DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)")

ben_mli_bfa_cmnn_wider <- ben_mli_bfa_cmnn %>% 
  pivot_wider(values_from = DALYs,
              names_from = Entity) %>% 
   mutate(across(starts_with("Benin") | starts_with("Burkina Faso") | starts_with("Mali"), ~round(., digits = 0)))

  kable(ben_mli_bfa_cmnn_wider)
```

## Plot Showing Trends in CMNN Burden Over Time

```{r echo = FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

ggplot(data = ben_mli_bfa_cmnn,
       aes(x = Year, y = DALYs, color = Entity)) +
  geom_line() +
  labs(title = "Line Graph for Benin, Burkina Faso, Mali",
       subtitle = "CMNN Burden, 1990 - 2020",
       x = "Year",
       y = "Value",
       color = "Country") +
  theme_minimal()
```

## Summary of CMNN Burden Findings

Overall, the DALYs for the CMNN in these countries seem to fall within the period under review.
This maybe due to effect of improved infrastructure or better healthcare over the years.
Benin seems to have a more slower and steady fall, perhaps due to a smaller population.
