# Non-Communicable Disease Burden (NCD)

```{r echo = FALSE, message = FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))
```

## Table of Estimates for NCD Burden Over Time

```{r echo = FALSE, message = FALSE, show_col_types = FALSE}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

ben_mli_bfa_ncd <- data_ncd %>%
  filter(Entity %in% c("Benin","Burkina Faso", "Mali")) %>% 
  select(Entity,Year,`DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  rename(DALYs = "DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)")

ben_mli_bfa_ncd_wider <- ben_mli_bfa_ncd %>% 
  pivot_wider(values_from = DALYs,
              names_from = Entity) %>% 
  mutate(across(starts_with("Benin") | starts_with("Burkina Faso") | starts_with("Mali"), ~round(., digits = 0)))  # Round the values


  kable(ben_mli_bfa_ncd_wider)
```

## Plot Showing Trends in NCD Burden Over Time 

```{r echo = FALSE, message = FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

ggplot(data = ben_mli_bfa_ncd,
       aes(x = Year, y = DALYs, color = Entity)) +
  geom_line() +
  labs(title = "Line Graph for Benin, Burkina Faso, Mali",
       subtitle = "NCD Burden, 1990 - 2020",
       x = "Year",
       y = "Value",
       color = "Country") +
  theme_minimal()
```

## Summary of NCD Burden Findings

The DALYs for NCD for the three countries begin to have a decline from 1990, but there is a rise in that of Burkina Faso from around 2010. This maybe due to some other internal factors such as change in government policies.
