# Overall Disease Burden

```{r echo = FALSE, message = FALSE}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the overall disease data
data_overall <- read_csv(here("data", "burden-of-disease-all-causes.csv"))

```

## Table of Estimates for Overall Disease Burden Over Time

```{r echo = FALSE, message = FALSE, show_col_types = FALSE}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

ben_mli_bfa_data_overall <- data_overall %>% 
  filter(Entity %in% c("Benin","Burkina Faso", "Mali")) %>% 
  select(Entity,Year,`DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`) %>% 
  rename(DALYs = "DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)")

ben_mli_bfa_data_overall_wider <- ben_mli_bfa_data_overall %>% 
  pivot_wider(values_from = DALYs,
              names_from = Entity) %>% 
  mutate(across(starts_with("Benin") | starts_with("Burkina Faso") | starts_with("Mali"), ~round(., digits = 0)))
  
kable(ben_mli_bfa_data_overall_wider)
```


## Plot Showing Trends in Overall Disease Burden Over Time

```{r echo = FALSE}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

ggplot(data = ben_mli_bfa_data_overall,
       aes(x = Year, y = DALYs, color = Entity)) +
  geom_line() +
  labs(title = "Line Graph for Benin, Burkina Faso, Mali",
       subtitle = "Overall Disease Burden, 1990 - 2020",
       x = "Year",
       y = "Value",
       color = "Country") +
  theme_minimal()
```
```

## Summary of Overall Disease Burden Findings

The DALYs for the overall disease burden for the three countries of Benin, Burkina Faso and Mali is seen to consistently decrease over time. Benin seems to have a sharper decline, compared to Burkina Faso and Mali, seemingly declining at the same pace. Perhaps political structure might and intervention programmes might have a role to play.
